Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2021-06-17 doc/manuscript/ESSDmanuscript.Rmd
update+knit

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'ESSD')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Sandro 
    surname: Federici 
    affiliation: "5" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "6" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "7" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Valentyna
    surname: Slivinska 
    affiliation: "5"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,6 *"
    email: teixeirak@si.edu
    corresponding: true

@@

>>>>

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'essd')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Sandro 
    surname: Federici 
    affiliation: "5" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "6" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "7" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Valentyna
    surname: Slivinska 
    affiliation: "5"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,6 *"
    email: teixeirak@si.edu
    corresponding: true

@@

### Biomass
Biomass includes living vegetation, above- and below-ground.

The IPCC defines aboveground biomass as "all biomass of living vegetation, both woody and herbaceous, above the soil including stems, stumps, branches, bark, seeds, and foliage" [].

Belowground biomass is defined as "all biomass of live roots" [].

### Dead Organic Matter
### Soil Organic Matter

### 2021-06-17 doc/manuscript/ESSDmanuscript.tex
update+knit

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
%
\usepackage{algorithmic}
\usepackage{algorithm}

@@

Forest ecosystem C stocks may be parsed into pools in various ways. IPCC
parses into biomass (aboveground and belowground), dead organic matter
(dead wood and litter), and soil organic matter (Table 1). Quantifying
these requires a one-time measurement.

\newpage

\textbf{Table 1: variables with definitions and measurement methods.}
Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).
(\emph{Currently adding this as a figure (generated from original draft)
because kableExtra doesn't seem to work in this template, and I can't
quickly get the template format to work. Table that we want here is
``figures\_tables/C\_pools.csv''})

\begin{figure}
\includegraphics[width=12cm]{figures_tables/C_pools_fig} \caption{Table 1}\label{fig:unnamed-chunk-1}
\end{figure}

\begin{table*}[t]
\caption{This is a start at table 1 using the template format. }
\begin{tabular}{l c c c r}
\tophline

pool & subpool & definition & major sources of estimate variation & IPCC guidance \\
\middlehline
biomass & aboveground & all biomass of living vegetation, both woody and herbaceous, above the soil & allometry, min dbh & acceptable to exclude understory  \\

\bottomhline
\end{tabular}
\belowtable{I don't know how to adjust so that it doesn't run off the page.}
\end{table*}

@@

To support export of data to EFDB, and to improve the overall quality of
the ForC database, we added \#\# increment variables, implemented some
modest restructuring, resolved duplicate records, and conducted quality
control. This section describes changes relative to ForC v2.0
\citep{anderson-teixeira_forc_2018}.

@@

We added \#\# increment variables to the set of named and defined
variables.
\emph{(https://github.com/forc-db/IPCC-EFDB-integration/issues/6)} These
are directly related to C stocks as previously defined in ForC, with
``\emph{delta.}'' added in front of the variable name.

@@

\begin{figure}[H]
\includegraphics[width=12cm]{figures_tables/ForC_updates_placeholder_fig} \caption{Table of changes to ForC fields (placeholder) }\label{fig:unnamed-chunk-3}
\end{figure}

>>>>

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
%%\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%
\usepackage{algorithmic}
\usepackage{algorithm}

@@

Forest ecosystem C stocks may be parsed into pools in various ways. IPCC
parses into biomass (aboveground and belowground), dead organic matter
(dead wood and litter), and soil organic matter
(Table~@ref(table\_variables)). Quantifying these requires a one-time
measurement.

\subsubsection{Biomass}

Biomass includes living vegetation, above- and below-ground.

The IPCC defines aboveground biomass as ``all biomass of living
vegetation, both woody and herbaceous, above the soil including stems,
stumps, branches, bark, seeds, and foliage'' {[}{]}.

Belowground biomass is defined as ``all biomass of live roots'' {[}{]}.

\subsubsection{Dead Organic Matter}

\subsubsection{Soil Organic Matter}

\begin{table}

\caption{\label{tab:table_variables}\textbf{Variables with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{pool} & \textbf{subpool} & \textbf{definition} & \textbf{major sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
biomass & aboveground & all biomass of living vegetation, both woody and herbaceous, above the soil & allometry, min dbh & acceptable to exclude understory\\
\hline
 & belowground & all biomass of live roots & allometry, min dbh, assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
dead organic matter & dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min dbh,  ... & default min dbh = 10cm, but may be chosen by country\\
\hline
 & litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & min dbh for dead wood , .. & \\
\hline
soils & soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & default sampling depth = 30cm, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

To support export of data to EFDB, and to improve the overall quality of
the ForC database, we added ten increment variables, implemented some
modest restructuring, resolved duplicate records, and conducted quality
control. This section describes changes relative to ForC v2.0
\citep{anderson-teixeira_forc_2018}.

@@

We added ten increment variables to the set of named and defined
variables (or 20, counting \_OM and \_C versions), which previously
included only one (aboveground biomass increment, \emph{delta.agb}).
\emph{(https://github.com/forc-db/IPCC-EFDB-integration/issues/6)} These
are directly related to C stocks as previously defined in ForC, with
``\emph{delta.}'' added in front of the variable name.

@@

\begin{figure}[H]
\includegraphics[width=12cm]{figures_tables/ForC_updates_placeholder_fig} \caption{Table of changes to ForC fields (placeholder) }\label{fig:unnamed-chunk-1}
\end{figure}
